Calendar-based graphics for visualising people's daily schedules

### 2017-06-29 main.Rmd
added a second para to the intro

---
title: "Calendar-based graphics for visualising people's daily schedules"
author:
  - familyname: Wang
    othernames: Earo
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : earo.wang@monash.edu
    correspondingauthor: true
  - familyname: Cook
    othernames: Di
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : dicook@monash.edu
  - familyname: Hyndman
    othernames: Rob
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : rob.hyndman@monash.edu
fontsize: 11pt
papersize: a4
fontfamily: mathpazo
blind: false
cover: true
wpnumber: no/yr
jelcodes: C10,C14,C22
bibliography: references.bib
biblio-style: authoryear-comp
abstract: "A brief summary of our ideas"
keywords: "blah, blah"
subparagraph: true
toc: false
output: 
  bookdown::pdf_document2:
    template: monashwp.tex
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    number_sections: yes
    citation_package: biblatex
---

```{r initial, echo = FALSE, cache = FALSE, results = 'hide'}
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE,
  fig.path = 'figure/', cache.path = 'cache/', fig.align = 'center', 
  fig.show = 'hold', cache = TRUE, external = TRUE, dev = "svglite")
```

@@

## Construction

@@

## Reference lines and grids

@@

## Extensions

The algorithm can be extended relatively easily to layout multiple years, or to lay out different calendar formats like weeks of a year. A longer term plan is to construct a new facet system for the grammar of graphics, extending the R package `ggplot2`, incorporating this structure, to enable for hierarchical faceting.

## Interactive graphics

We shall enable interactivity in the calendar-based graphics for time series data. It will allow users to transform different temporal components and switch displays between overlaying and faceting through key strokes or mouse clicks.

## Application

@@

>>>>

---
title: "Calendar-based graphics for visualising people's daily schedules"
author:
  - familyname: Wang
    othernames: Earo
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : earo.wang@monash.edu
    correspondingauthor: true
  - familyname: Cook
    othernames: Di
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : dicook@monash.edu
  - familyname: Hyndman
    othernames: Rob J
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : rob.hyndman@monash.edu
fontsize: 11pt
papersize: a4
fontfamily: mathpazo
blind: false
cover: true
wpnumber: no/yr
jelcodes: C10,C14,C22
bibliography: references.bib
biblio-style: authoryear-comp
abstract: "A brief summary of our ideas"
keywords: "blah, blah"
subparagraph: true
toc: false
output: 
  bookdown::pdf_document2:
    template: monashwp.tex
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    number_sections: yes
    citation_package: biblatex
---

```{r initial, echo = FALSE, cache = FALSE, results = 'hide'}
library(knitr)
opts_chunk$set(
  warning = FALSE, message = FALSE, echo = FALSE,
  fig.path = 'figure/', cache.path = 'cache/', fig.align = 'center', 
  fig.show = 'hold', cache = TRUE, external = TRUE, dev = "svglite"
)
```

@@

Calendar-based visualisation proves useful in unfolding human-related activities in temporal context. For example, @van_wijk_cluster_1999 developed a calendar view of heatmap to represent the number of employees in the work place over a year, where colours indicate different clusters derived from the days. It effectively contrasts weekdays and weekends, highlights public holiday, and presents other known seasons like school vacations, all of which have influence over the turn-outs in the office. However, their technique is too constrained to colour-encoding graphics. We shall extend this calendar-based arrangement using linear algebra tools from a single type of heatmap to a broader range of glyphs, from univariate to multivariate cases, from one type of calendar to three types. The proposed algorithm has been implemented into the `sugrrants` package [@sugrrants] using R [@r].

This work is partly motivated by studying hourly foot traffic in the city of Melbourne [@ped]. There have been 43 sensors installed across downtown Melbourne since 2009 in order to capture the pulse of city life. (FIGURE ?: weekday at multiple sensors) Figure (?) lends itself to addressing some of the issues: (1) to simultaneously exhibit multiple seasons including time of day, day of week, and day of year, (2) to visualise long historical time series data in an information-dense way at the overview level, (3) to compare and contrast multiple time series on a common scale, (4) to quickly look up the date when there are unexpected events happening without interactivity.

# Construction

(NEED A BRIEF INTRO HERE)

## Calendar grids

@@

The algorithm can be extended relatively easily to layout multiple years, or to lay out different calendar formats like weeks of a year.

## Scales

## Reference lines and labels

@@

(INTERACTIVITY COULD BE DISCUSSED HERE) We shall enable interactivity in the calendar-based graphics for time series data. It will allow users to transform different temporal components and switch displays between overlaying and faceting through key strokes or mouse clicks.

# Applications

@@

# Conclusions

The limitation is also evident: hard to perceive trend as not on the common scale.
